SEM with site-level algae data

### 2025-01-24 SEM_quad.Rmd

@@

```{r}
# quad-level data summarized with one point for all time
algae <- read_csv(here("data/diversity_stability_synchrony.csv"))
```

@@

```{r}
algae %>% 
  mutate(
    richness_mean_scale = scale(algae$richness_mean)[,1],
    cover_mean_scale = scale(algae$cover_mean)[,1],
    cover_stability_scale = scale(algae$cover_stability)[,1],
    synchrony_scale = scale(algae$synchrony)[,1]
  ) -> algae_scaled
```

# Make a simple SEM

@@

```{r}
# effect of richness on synchrony
m.sprich.synch <- glmer(synchrony_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_scaled)
  performance::check_model(m.sprich.synch)

# effect of synchrony on stability
m.synch.stab.cov <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)  
  performance::check_model(m.synch.stab)

@@

```{r}
# Build SEM
sem <- psem(
  m.sprich.synch,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae)
)

@@

```{r}
# Extract coefficients
#coefficients <- summary(sem)$coefficients

# Add a column for color based on the sign of the path coefficient
#coefficients$color <- ifelse(coefficients$Estimate > 0, "blue", #"red")

plot(sem)
```

@@

```{r}
# effect of richness on synchrony
m.sprich.synch.FR <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Fringing", ])
  performance::check_model(m.sprich.synch.FR)

# effect of synchrony on stability
m.synch.stab.cov.FR <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Fringing", ])  
  performance::check_model(m.synch.stab.cov.FR)

@@

```{r}
# Build SEM
sem_fringe_brac <- psem(
  m.sprich.synch.FR,
  m.synch.stab.cov.FR,
  m.sprich.cov.FR,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Fringing", ])
)

@@

```{r}
# effect of richness on synchrony
m.sprich.synch.BA <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Backreef", ])
  performance::check_model(m.sprich.synch.BA)

# effect of synchrony on stability
m.synch.stab.cov.BA <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Backreef", ])  
  performance::check_model(m.synch.stab.cov.BA)

@@

```{r}
# Build SEM
sem_back_brac <- psem(
  m.sprich.synch.BA,
  m.synch.stab.cov.BA,
  m.sprich.cov.BA,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Backreef", ])
)

@@

```{r}
# effect of richness on synchrony
m.sprich.synch.FO10 <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
  performance::check_model(m.sprich.synch.FO10)

# effect of synchrony on stability
m.synch.stab.cov.FO10 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])  
  performance::check_model(m.synch.stab.cov.FO10)

@@

```{r}
# Build SEM
sem_for10_brac <- psem(
  m.sprich.synch.FO10,
  m.synch.stab.cov.FO10,
  m.sprich.cov.FO10,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
)

@@

```{r}
# effect of richness on synchrony
m.sprich.synch.FO17 <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
  performance::check_model(m.sprich.synch.FO17)

# effect of synchrony on stability
m.synch.stab.cov.FO17 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])  
  performance::check_model(m.synch.stab.cov.FO17)

@@

```{r}
# Build SEM
sem_for17_brac <- psem(
  m.sprich.synch.FO17,
  m.synch.stab.cov.FO17,
  m.sprich.cov.FO17,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
)

>>>>

# clear workspace
rm(list = ls())

@@

```{r}
# functional richness from Noam
functrich <- read_csv(here("data/full_plot_level_dss.csv"))

# quad-level data summarized with one point for all time
algae_1 <- read_csv(here("data/diversity_stability_synchrony.csv"))
```

These two guys have different dimensions. See where they differ:

```{r}
functrich[!(functrich$location %in% algae_1$location),] 
```

Noam says we can just inner join/exclude these--just used to calculate averages

```{r}
functrich %>% 
  select(location, habitat, site, functional_richness_mean) %>% 
  inner_join(algae_1) -> algae
```

@@

```{r}
algae %>% 
  mutate(
    richness_mean_scale = scale(algae$richness_mean)[,1],
    functional_richness_mean_scale = scale(algae$functional_richness_mean)[,1],
    cover_mean_scale = scale(algae$cover_mean)[,1],
    cover_stability_scale = scale(algae$cover_stability)[,1],
    synchrony_scale = scale(algae$synchrony)[,1]
  ) -> algae_scaled
```

# Make an overall SEM with functional richness

@@

```{r}
# effect of richness on synchrony
m.sprich.synch <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_scaled)
  performance::check_model(m.sprich.synch)

# effect of taxonomic richness on functional richness
m.sprich.funrich <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_scaled)
  performance::check_model(m.sprich.funrich)

# effect of synchrony on stability
m.synch.stab.cov <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)  
  performance::check_model(m.synch.stab.cov)

@@

```{r}
# Build SEM
sem <- psem(
  m.sprich.synch,
  m.sprich.funrich,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae_scaled)
)

@@

```{r}
plot(sem)
```

# Make an overall SEM withOUT functional richness

```{r}
# effect of richness on synchrony
m.sprich.synch.noFR <- glmer(synchrony_scale ~ richness_mean_scale  + (1|site) + (1|habitat),
                        data = algae_scaled)
  performance::check_model(m.sprich.synch.noFR)

# effect of synchrony on stability
m.synch.stab.cov.noFR <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)  
  performance::check_model(m.synch.stab.cov.noFR)

# effect of richness on cover
m.sprich.cov.noFR <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)
  performance::check_model(m.sprich.cov.noFR)
```

```{r}
# Build SEM
sem.noFR <- psem(
  m.sprich.synch,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae_scaled)
)

summary(sem.noFR)
```

```{r}
plot(sem.noFR)
```

@@

```{r}
# effect of richness on synchrony
m.sprich.synch.FR <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Fringing", ])
  performance::check_model(m.sprich.synch.FR)

# effect of taxonomic richness on functional richness
m.sprich.funrich.FR <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Fringing", ])
  performance::check_model(m.sprich.funrich.FR)

# effect of synchrony on stability
m.synch.stab.cov.FR <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale +  cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Fringing", ])  
  performance::check_model(m.synch.stab.cov.FR)

@@

```{r}
# Build SEM
sem_fringe_brac <- psem(
  m.sprich.synch.FR,
  m.sprich.funrich.FR,
  m.synch.stab.cov.FR,
  m.sprich.cov.FR,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Fringing", ])
)

@@

```{r}
# effect of richness on synchrony
m.sprich.synch.BA <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Backreef", ])
  performance::check_model(m.sprich.synch.BA)

# effect of taxonomic richness on functional richness
m.sprich.funrich.BA <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Backreef", ])
  performance::check_model(m.sprich.funrich.BA)

# effect of synchrony on stability
m.synch.stab.cov.BA <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + functional_richness_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Backreef", ])  
  performance::check_model(m.synch.stab.cov.BA)

@@

```{r}
# Build SEM
sem_back_brac <- psem(
  m.sprich.synch.BA,
  m.sprich.funrich.BA,
  m.synch.stab.cov.BA,
  m.sprich.cov.BA,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Backreef", ])
)

@@

```{r}
# effect of richness on synchrony
m.sprich.synch.FO10 <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
  performance::check_model(m.sprich.synch.FO10)

# effect of taxonomic richness on functional richness
m.sprich.funrich.FO10 <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
  performance::check_model(m.sprich.funrich.FO10)

# effect of synchrony on stability
m.synch.stab.cov.FO10 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + functional_richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])  
  performance::check_model(m.synch.stab.cov.FO10)

@@

```{r}
# Build SEM
sem_for10_brac <- psem(
  m.sprich.synch.FO10,
  m.sprich.funrich.FO10,
  m.synch.stab.cov.FO10,
  m.sprich.cov.FO10,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
)

@@

```{r}
# effect of richness on synchrony
m.sprich.synch.FO17 <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
  performance::check_model(m.sprich.synch.FO17)

# effect of taxonomic richness on functional richness
m.sprich.funrich.FO17 <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
  performance::check_model(m.sprich.funrich.FO17)

# effect of synchrony on stability
m.synch.stab.cov.FO17 <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])  
  performance::check_model(m.synch.stab.cov.FO17)

@@

```{r}
# Build SEM
sem_for17_brac <- psem(
  m.sprich.synch.FO17,
  m.sprich.funrich.FO17,
  m.synch.stab.cov.FO17,
  m.sprich.cov.FO17,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
)

### 2025-01-24 SEM_site.Rmd

# Make a simple SEM

@@

```{r}
# effect of richness on synchrony
m.sprich.synch <- glmer(synchrony_mean_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_site_scaled)
  performance::check_model(m.sprich.synch)

@@

```{r}
# Build SEM
sem <- psem(
  m.sprich.synch,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae_site_scaled)
)

@@

>>>>

# Make an overall SEM

@@

```{r}
# effect of richness on synchrony
m.sprich.synch <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_site_scaled)
  performance::check_model(m.sprich.synch)

# effect of taxonomic richness on functional richness
m.sprich.funrich <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_site_scaled)
  performance::check_model(m.sprich.funrich)

@@

```{r}
# Build SEM
sem <- psem(
  m.sprich.synch,
  m.synch.stab.cov,
  m.sprich.cov,
  m.sprich.funrich,
  data = as.data.frame(algae_site_scaled)
)

@@

Poor convergence issues/singularity

# What about no random effect for habitat?

```{r}
# effect of richness on synchrony
m.sprich.synch.nR <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_site_scaled)
  performance::check_model(m.sprich.synch.nR)

# effect of taxonomic richness on functional richness
m.sprich.funrich.nR <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_site_scaled)
  performance::check_model(m.sprich.funrich.nR)

# effect of synchrony on stability
m.synch.stab.cov.nR <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale + spatial_synchrony_scale + (1|site),
                      data = algae_site_scaled)  
  performance::check_model(m.synch.stab.cov.nR)

# effect of richness on cover
m.sprich.cov.nR <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_site_scaled)
  performance::check_model(m.sprich.cov.nR)
```

```{r}
# Build SEM
sem.nR <- psem(
  m.sprich.synch.nR,
  m.synch.stab.cov.nR,
  m.sprich.cov.nR,
  m.sprich.funrich.nR,
  data = as.data.frame(algae_site_scaled)
)

summary(sem.nR)
```

```{r}
plot(sem.nR)
```
